---
title: Data structures
layout: default
output: html_document
---

```{r include=FALSE}
tutorial::go_interactive()
```

# Data structures {#data-structures}

This chapter summarises the most important data structures in base R. You've probably used many (if not all) of them before, but you may not have thought deeply about how they are interrelated. In this brief overview, I won't discuss individual types in depth. Instead, I'll show you how they fit together as a whole. If you need more details, you can find them in R's documentation.

R's base data structures can be organised by their dimensionality (1d, 2d, or nd) and whether they're homogeneous (all contents must be of the same type) or heterogeneous (the contents can be of different types). This gives rise to the five data types most often used in data analysis: 

|    | Homogeneous   | Heterogeneous |
|----|---------------|---------------|
| 1d | Atomic vector | List          |
| 2d | Matrix        | Data frame    |
| nd | Array         |               |

Almost all other objects are built upon these foundations. In [the OO field guide](#oo) you'll see how more complicated objects are built of these simple pieces. Note that R has no 0-dimensional, or scalar types. Individual numbers or strings, which you might think would be scalars, are actually vectors of length one. 

Given an object, the best way to understand what data structures it's composed of is to use `str()`. `str()` is short for structure and it gives a compact, human readable description of any R data structure. \indexc{str()}

## Vectors {#vectors}

The basic data structure in R is the vector. Vectors come in two flavours: atomic vectors and lists. They have three common properties:

* Type, `typeof()`, what it is.
* Length, `length()`, how many elements it contains.
* Attributes, `attributes()`, additional arbitrary metadata.

They differ in the types of their elements: all elements of an atomic vector must be the same type, whereas the elements of a list can have different types.

NB: `is.vector()` does not test if an object is a vector. Instead it returns `TRUE` only if the object is a vector with no attributes apart from names. Use `is.atomic(x) || is.list(x)` to test if an object is actually a vector.

### Atomic vectors

There are four common types of atomic vectors that I'll discuss in detail: logical, integer, double (often called numeric), and character. There are two rare types that I will not discuss further: complex and raw. \index{atomic vectors} \index{vectors!atomic|see{atomic vectors}}

Atomic vectors are usually created with `c()`, short for combine: \indexc{c()}

```{r}
dbl_var <- c(1, 2.5, 4.5)
# With the L suffix, you get an integer rather than a double
int_var <- c(1L, 6L, 10L)
# Use TRUE and FALSE (or T and F) to create logical vectors
log_var <- c(TRUE, FALSE, T, F)
chr_var <- c("these are", "some strings")
```

Atomic vectors are always flat, even if you nest `c()`'s:

```{r}
c(1, c(2, c(3, 4)))
# the same as
c(1, 2, 3, 4)
```

Missing values are specified with `NA`, which is a logical vector of length 1. `NA` will always be coerced to the correct type if used inside `c()`, or you can create `NA`s of a specific type with `NA_real_` (a double vector), `NA_integer_` and `NA_character_`. \indexc{NA}

#### Types and tests

Given a vector, you can determine its type with `typeof()`, or check if it's a specific type with an "is" function: `is.character()`, `is.double()`, `is.integer()`, `is.logical()`, or, more generally, `is.atomic()`. \indexc{typeof()}

```{r}
int_var <- c(1L, 6L, 10L)
typeof(int_var)
is.integer(int_var)
is.atomic(int_var)

dbl_var <- c(1, 2.5, 4.5)
typeof(dbl_var)
is.double(dbl_var)
is.atomic(dbl_var)
```

NB: `is.numeric()` is a general test for the "numberliness" of a vector and returns `TRUE` for both integer and double vectors. It is not a specific test for double vectors, which are often called numeric. \indexc{is.numeric()}

```{r}
is.numeric(int_var)
is.numeric(dbl_var)
```
